---
title: "Merge by PRIO"
output: html_document
date: "2024-04-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Aggregate by PRIO_GRID

```{r}
library(tidyr)
library(tidyverse)
library(here)
library(haven)

# Import the data
grd <- read_dta(here("Data", "GRD Dataset", "dfhsw_GRD_public_v1.dta"))

grd |> 
  head()

```


## Filter for country and data range
Now let's filter for world and the year range matching the conflict dataset. We also drop some irrelevant columns. I include 1994 to form the baseline lootability estimate (before sample period starting 1995). 1994 is the earliest year in the dataset for world.

```{r}
# Filter for world
world_grd <- grd |> 
  filter(year >= 1994 & year <= 2014)

world_grd |>
  head()
```
```{r}
# First display how many unique values there are of gid
unique_prio <- world_grd |> 
  select(gid) |> 
  distinct() |> 
  nrow()

print(paste("Unique prio values:", unique_prio))

# Print the names of gid unique values
unique_prio_names <- world_grd |> 
  select(gid) |> 
  distinct() |> 
  pull()

print(unique_prio_names)
```
```{r}
# Check for duplicates
world_grd |>
  select(-obs_no) |>
  duplicated() |>
  sum()

```

```{r}
# Step 1: Add an identifier for each group of duplicates (except for 'obs_no')
resources_marked <- world_grd %>%
  group_by(across(-obs_no)) %>% 
  mutate(dup_id = row_number()) %>%
  ungroup()

# Step 2: Keep only the first of each set of duplicates
world_filtered <- resources_marked %>%
  filter(dup_id == 1) %>%
  select(-dup_id)  # Removing the temporary 'dup_id' column

# Optional: Check the new size of the dataset
print(paste("New dataset size:", nrow(world_filtered)))
```

# Aggregate prio-year pairs to one row
```{r}
# Count the number of PRIO_GID-year pairs for each resource type
duplicates_count <- world_filtered %>%
  group_by(gid, year, resource) %>%
  summarise(count = n(), .groups = 'drop') %>%
  filter(count > 1)

# Display the counts of duplicates
print(duplicates_count)

# Optionally, to see the total number of duplicates
total_duplicates <- sum(duplicates_count$count) - nrow(duplicates_count)
print(paste("Total number of duplicate observations:", total_duplicates))
```
We should expect to remove 18638 observations

```{r}
# Define lists of columns based on their aggregation rule
sum_columns <- c("annuallocationcapacity", "comtrade_value", "wb_value", "usgs_value", "multicolour_value",
                 "world_val_nomc", "world_val_withmc", "wd_annual_value_location1", "wd_annual_value_location2",
                 "exp_annual_value_location1", "exp_annual_value_location2")

# Define a custom summarise function for non-sum columns
concat_if_different <- function(x) {
  unique_x <- unique(x)
  if(length(unique_x) == 1) {
    return(as.character(unique_x))
  } else {
    return(paste(unique_x, collapse = "; "))
  }
}

world_aggregated <- world_filtered %>%
  group_by(gid, year, resource) %>%
  summarise(across(all_of(sum_columns), sum, na.rm = TRUE), # Sum the defined columns
            across(-all_of(sum_columns), concat_if_different), # Concatenate if different for the rest
            .groups = 'drop') # Prevents the result from being grouped

```

## Verify same-resource aggregation
Let's check if the aggregation was successful.

```{r}
# Check if the aggregation was successful
duplicates_after_aggregation <- world_aggregated %>%
  group_by(gid, year, resource) %>%
  summarise(count = n(), .groups = 'drop') %>%
  filter(count > 1)

# Check if there are any duplicates left
if (nrow(duplicates_after_aggregation) > 0) {
  print("There are remaining duplicates after aggregation.")
} else {
  print("No duplicates remain after aggregation. Aggregation was successful.")
}

```
# Aggregate prio-year pairs to one row (different resource types)
Now we aggregate the rows with the same gid and year but different resources. We will keep the resource-invariant variables as they are and create new variables for the resource-variant variables by reshaping the data.


Now let's aggregate the data. 
```{r}
columns_to_reshape <- setdiff(names(world_aggregated), c("gid", "year"))

# Pivot the dataset to wide format
world_wide <- world_aggregated %>%
  pivot_wider(names_from = resource, names_prefix = "res_", 
              values_from = all_of(columns_to_reshape),
              values_fill = list(value = NA)) # Fill missing values with NA

# Note: The values_fill argument is set to fill with NA for simplicity, 
# adjust as needed based on your specific requirements

# Print the result to check
print(world_wide)

# Count the final number of columns to verify
final_num_columns <- ncol(world_wide)
print(paste("Final number of columns:", final_num_columns))
```


```{r}
# --- Step 1: Verify Initial Conditions Before Reshaping ---
# Count unique gid-year combinations in the aggregated data
unique_gid_year_combinations <- nrow(distinct(world_aggregated, gid, year))

# Calculate the expected number of columns after reshaping
num_unique_resources <- length(unique(world_aggregated$resource))
columns_to_reshape <- setdiff(names(world_aggregated), c("gid", "year"))
expected_columns_after_reshaping <- 2 + (num_unique_resources * length(columns_to_reshape))

# --- Reshaping (already provided) ---
world_wide <- world_aggregated %>%
  pivot_wider(names_from = resource, names_prefix = "res_", 
              values_from = all_of(columns_to_reshape),
              values_fill = list(value = NA))

# --- Step 3: Verification after Reshaping ---
# Verify number of rows matches unique gid-year combinations
actual_rows_reshaped <- nrow(world_wide)
if (actual_rows_reshaped == unique_gid_year_combinations) {
  print("Rows verification passed: Each observation has a unique gid-year pair.")
} else {
  print("Rows verification failed: Mismatch in expected and actual number of unique gid-year pairs.")
}

# Verify the number of columns
actual_columns_reshaped <- ncol(world_wide)
if (actual_columns_reshaped == expected_columns_after_reshaping) {
  print("Columns verification passed: The number of columns matches the expected number after reshaping.")
} else {
  print(paste("Columns verification failed: Expected", expected_columns_after_reshaping, "columns, but found", actual_columns_reshaped, "columns."))
}

# Check for uniqueness of gid-year pairs in the reshaped dataset
unique_pairs_post_reshape <- world_wide %>%
  distinct(gid, year) %>%
  nrow()
if (unique_pairs_post_reshape == actual_rows_reshaped) {
  print("Post-reshaping verification passed: Each gid-year pair is unique in the reshaped dataset.")
} else {
  print("Post-reshaping verification failed: Some gid-year pairs may not be unique.")
}

```

Create lootability measure

```{r}
library(dplyr)

# Find all lootable flag columns
lootable_flags <- names(world_wide)[grepl("^lootable_", names(world_wide))]

# Assuming lootable_flags are defined, as well as the corresponding world_value_columns
world_value_columns <- gsub("lootable_", "wd_annual_value_location1_", lootable_flags)

# Convert the data frame to a tibble for better handling in dplyr
world_wide <- as_tibble(world_wide)

# Create a new dataframe to store the results
aggregate_lootability <- tibble(gid = character(), year = integer(), total_lootable_value = numeric())

# Loop through each lootable resource and perform aggregation
for (i in seq_along(lootable_flags)) {
    lootable_col <- lootable_flags[i]
    value_col <- world_value_columns[i]

    # Check if the current value column exists in the dataframe
    if (value_col %in% names(world_wide)) {
        # Filter and aggregate using dplyr
        temp_agg <- world_wide %>%
            filter(!!sym(lootable_col) == 1) %>%  # Filter for lootable resources
            group_by(gid, year) %>%
            summarise(
                lootable_value = sum(!!sym(value_col), na.rm = TRUE),
                .groups = 'drop'
            ) %>%
            rename(total_lootable_value = lootable_value)  # Rename to a common column for easier merging

        # Merge or bind rows with the main aggregate dataframe
        if (nrow(aggregate_lootability) == 0) {
            aggregate_lootability <- temp_agg
        } else {
            # If already initialized, merge the new data
            aggregate_lootability <- full_join(aggregate_lootability, temp_agg, by = c("gid", "year")) %>%
                rowwise() %>%
                mutate(total_lootable_value = sum(c_across(starts_with("total_lootable_value")), na.rm = TRUE)) %>%
                select(gid, year, total_lootable_value)
        }
    }
}

# Join the aggregated lootability data back to the main dataframe
world_wide <- left_join(world_wide, aggregate_lootability, by = c("gid", "year"))

# Replace NA values with zero in the lootability column
world_wide$total_lootable_value <- replace_na(world_wide$total_lootable_value, 0)

# Check the results
print(head(world_wide))


```


Now, normalise lootability measure by dividing by total value of production of all resources in that year. Calculate lootability measure with world_value_location1 columns

```{r}
world_wide <- world_wide %>%
  # Convert to tibble if not already
  as_tibble() %>%
  # Calculate the total value of all world value columns
  mutate(total_value = rowSums(select(., all_of(world_value_columns)), na.rm = TRUE)) %>%
  # Use the previously created total_lootable_value to calculate lootability proportion
  mutate(lootability_proportion = case_when(
    total_value > 0 ~ total_lootable_value / total_value,
    TRUE ~ 0  # Avoid division by zero by setting proportion to 0 when total_value is 0
  ))

# Print head
print(head(world_wide))
```
Let's save it as a csv and dta file

```{r}
# Save as CSV
write_csv(world_wide, here("Data", "GRD Dataset", "world_wide_data.csv"))
```

```{r}
# Save total_value, total_lootable_value and lootability_proportion as a csv and dta file
world_lootablecols <- world_wide %>%
  select(gid, year, total_value, total_lootable_value, lootability_proportion)

# Save as CSV
write_csv(world_lootablecols, here("Data", "GRD Dataset", "world_lootablecols.csv"))

# Save as DTA
write_dta(world_lootablecols, here("Data", "GRD Dataset", "world_lootablecols.dta"))

```

# Irrelevant after here
```{r}
# See how many gid cells have non-zero lootability_proportion
library(dplyr)

count_positive_gids <- world_wide%>%
  filter(lootability_proportion > 0) %>%    # Keep only rows where lootable is positive
  group_by(gid) %>%           # Group by 'gid'
  summarise(has_positive = any(lootability_proportion > 0), .groups = 'drop') %>% # Check if there are any positive values
  summarise(count = n())      # Count the number of gids with positive values

print(count_positive_gids)

```
It has 1122 different gid cells with positive lootability_proportion. The denly set has 1302 with their naive measure of lootable.

# Merging

Now let's merge with the conflict dataset keeping all columns from both datasets.

```{r}
# Import conflict data
conflict_data <- read_csv(here("Data", "xSub Conflict Data", "xSub_MELTT1km1dB_COL_priogrid_year.csv"))
```
```{r}
# Rename the column in conflict data to match the admin1 column in resource data
conflict_data <- conflict_data |> 
  rename(gid = PRIO_GID) |> 
  rename(year = YEAR) |> 
  filter(year >= 1994 & year <= 2014)
```

Check whether they overlap and how many should be merged

```{r}
library(dplyr)

# Assuming 'gid' columns are properly named and present in both datasets
# If the column names are different, adjust them accordingly before this step

# Extract gid values from both datasets
gids_conflict <- unique(conflict_data$gid)
gids_resource <- unique(world_wide$gid)

# Find common gids (intersection)
common_gids <- intersect(gids_conflict, gids_resource)

# Find gids unique to conflict_data
unique_conflict_gids <- setdiff(gids_conflict, gids_resource)

# Find gids unique to world_wide
unique_resource_gids <- setdiff(gids_resource, gids_conflict)

# Print the results
cat("Common GIDs:", length(common_gids), "\n")
cat("GIDs unique to Conflict Data:", length(unique_conflict_gids), "\n")
cat("GIDs unique to Resource Data:", length(unique_resource_gids), "\n")

```

```{r}
merged_data <- full_join(conflict_data, world_wide, by = c("gid", "year"))
```

Check the merge was succesful

```{r}
world_gid_rows <- merged_data %>%
  filter(gid %in% unique(world_wide$gid))  # Assuming 'gid' is the column name in both datasets

# Display the rows
print(world_gid_rows)
```



```{r}
# Save data
write_csv(merged_data, here("Data", "Merged Data", "merged_data_COL_prio_year.csv"))
```


Merge keeping only mining data:

```{r}
merged_data_left <- left_join(world_wide, conflict_data, by = c("gid", "year"))

# Check the results to ensure the merge worked as expected
head(merged_data_left)

```


# Trying some initial regressions
```{r}
library(tidyverse)

# Load your dataset if not already loaded
# df <- read.csv("your_data.csv")

# Assuming 'df' is your dataframe, and it contains 'ACTION_ANY' and 'lootability_prop'
# Fit a Poisson regression model
poisson_model <- glm(ACTION_ANY ~ lootability_proportion, family = poisson(), data = merged_data)

# Display the summary of the model to see results
summary(poisson_model)

```
No 


Check how many gids have of non-zero resource values for each resource.

```{r}
# Assuming 'merged_data' is your dataframe
# Identify columns that match the pattern for resource capacities
resource_columns <- names(merged_data)[grepl("^annuallocationcapacity_res_", names(merged_data))]

# Initialize a data frame to store results
results <- data.frame(resource = character(), unique_gids = integer())

# Process each resource column to count non-zero, unique gid values
for (resource in resource_columns) {
  unique_gids_count <- merged_data %>%
    filter(.data[[resource]] != 0) %>%  # Filter non-zero entries for the current resource
    distinct(gid) %>%                  # Get distinct gids with non-zero entries
    nrow()                             # Count unique gids

  # Append the results for the current resource to the results data frame
  results <- rbind(results, data.frame(resource = resource, unique_gids = unique_gids_count))
}

# Display the results for all resources
print(results)
```



Save merged data as dta file
Rename these variables to shorter names:
annuallocationcapacity_res_sulfur`, `annuallocationcapacity_res_phosphate`,
  `annuallocationcapacity_res_black carbon`, `annuallocationcapacity_res_iron and steel`,
  `annuallocationcapacity_res_cement`, `annuallocationcapacity_res_kaolin`,
  `annuallocationcapacity_res_platinum`, `annuallocationcapacity_res_emerald`,
  `annuallocationcapacity_res_silver`, `annuallocationcapacity_res_copper`,
  `annuallocationcapacity_res_natural gas`, `annuallocationcapacity_res_nickel`,
  `annuallocationcapacity_res_nitrogen`, `comtrade_value_res_iron and steel`,
  `multicolour_value_res_black carbon`, `multicolour_value_res_iron and steel`,
  `multicolour_value_res_natural gas`, `world_val_nomc_res_iron and steel`, …,
  `log_multicolour_val_res_iron and steel`, and `log_multicolour_val_res_natural gas

```{r}
# Extract the original variable names
original_vars <- names(merged_data)

# Function to clean and rename variable names
clean_var_names <- function(var_name) {
  # Pattern replacement for specific types of names
  if (grepl("^resource_res_", var_name)) {
    var_name <- sub("^resource_res_", "", var_name)
  } else if (grepl("^annuallocationcapacity_res_", var_name)) {
    var_name <- sub("^annuallocationcapacity_res_", "ann_cap_", var_name)
  } else if (grepl("^comtrade_value_res_", var_name)) {
    var_name <- sub("^comtrade_value_res_", "UN_val_", var_name)
  } else if (grepl("^wb_value_res_", var_name)) {
    var_name <- sub("^wb_value_res_", "wb_val_", var_name)
  } else if (grepl("^export_price_first_mult1_res_", var_name)) {
    var_name <- sub("^export_price_first_mult1_res_", "expp1_mult1_", var_name)
  }
  
  # Replace spaces and special characters
  var_name <- gsub(" ", "_", var_name)      # Replace spaces with '_'
  var_name <- gsub(" and ", "_", var_name)  # Replace ' and ' with '_'
  
  # Ensure name does not start with a number or underscore
  if (grepl("^[0-9]|^_", var_name)) {
    var_name <- paste0("var_", var_name)
  }
  # Truncate names to 32 characters
  substring(var_name, 1, 32)
}

# Apply the cleaning function to variable names
names(merged_data) <- sapply(original_vars, clean_var_names)



```


```{r}
# Assume 'merged_data' is your main dataframe and it includes 'year' and 'gid' columns
# Filter data for the year 1994 and select the 'ann_cap_oil' column
oil_1994_data <- merged_data[merged_data$year == 1994, c("gid", "ann_cap_oil")]

# Remove any duplicates if 'gid' isn't unique for the year 1994
oil_1994_data <- oil_1994_data[!duplicated(oil_1994_data$gid), ]

# Merge this back to the original dataframe on 'gid' to create the new variable
merged_data <- merge(merged_data, oil_1994_data, by = "gid", all.x = TRUE, suffixes = c("", "_1994"))

# Rename the merged 'ann_cap_oil' column to 'oil_prod_1994'
names(merged_data)[names(merged_data) == "ann_cap_oil_1994"] <- "oil_prod_1994"

# Check the first few rows to confirm the new column has been added
head(merged_data)

```
```{r}
# Ensure all variables are of type double
merged_data[numeric_vars] <- lapply(merged_data[numeric_vars], function(x) as.numeric(as.character(x)))

# Recheck numeric variables in case of previous type issues
numeric_vars <- sapply(merged_data, is.numeric)

# Identify variables that match the pattern 'expp1_mult1_*'
price_vars <- grep("expp1_mult1_", names(merged_data), value = TRUE)

# Filter price_vars to ensure they are also numeric after conversions
price_vars <- price_vars[price_vars %in% names(merged_data[numeric_vars])]

# Apply log-plus-one transformation to these variables
for (var in price_vars) {
  # Create new variable name for the log-transformed variable
  new_var_name <- paste("log", var, sep="_")
  
  # Check if the variable is indeed numeric; convert if not already (as a precaution)
  if (!is.numeric(merged_data[[var]])) {
    merged_data[[var]] <- as.numeric(merged_data[[var]])
  }

  # Calculate the log of the variable plus one, ensuring NA values remain as NA
  merged_data[[new_var_name]] <- ifelse(is.na(merged_data[[var]]), 
                                        NA, 
                                        log(merged_data[[var]] + 1))
}

# Check and handle any further issues if this still causes errors

```


```{r}
# Save the merged data to a Stata .dta file
library(haven)
write_dta(merged_data, here("Data", "Merged Data", "merged_data_COL_prio_year.dta"))
```


```{r}
print(names(merged_data))
```

```{r}
rm(list = ls())
library(tidyr)
library(tidyverse)
library(here)
library(haven)
world_wide <- read_csv(here("Data", "GRD Dataset", "world_wide_data.csv"))
```
# Mine Ownership Score
```{r}
# check if there are any variables with the name mineownership in it
mineownership_vars <- grep("mineownership", names(world_wide), value = TRUE)

# Print head for these columns
owner <- world_wide %>%
  select(all_of(mineownership_vars))

  
```

```{r}
# IMport grd
grd <- read_dta(here("Data", "GRD Dataset", "dfhsw_GRD_public_v1.dta"))
```

```{r}
library(knitr)
library(kableExtra)

# Count occurrences of each unique value in the mineownership column
mineownership_counts <- grd %>%
  count(mineownership, sort = TRUE) %>%
  rename(`Mine Ownership` = mineownership, `Occurrences` = n)

# Print distinct values (for verification, can be removed in final script)
print(mineown_distinct)

# Create a kable table with caption and label
kable_table <- mineownership_counts %>%
  kable("latex", booktabs = TRUE, caption = "Occurrences of Each Mine Ownership Type", label = "tab:mineownership_counts") %>%
  kable_styling(latex_options = c("hold_position", "striped"))

# Export the table to a .tex file
output_path <- here("mineownership_counts_table.tex")
writeLines(kable_table, output_path)
```


```{r}
# Define the ownership score mapping based on the given entries and classification
ownership_scores <- c(
  "industrial" = 3,                    # Domestic Private
  "government" = 4,                    # Public
  "n/a" = 0,                           # No information
  "government and industrial" = 3.5,   # Mixed categories (shared ownership)
  "artisanal" = 3,                     # Domestic Private
  "shared" = 2.5,                      # Mixed categories (shared ownership)
  "foreign government" = 3.5,          # Foreign
  "cooperative and industrial" = 2.5,  # Mixed categories (shared ownership)
  "artisanal, military, rebel" = 1.5,  # Rebel/Military involvement
  "rebel" = 1.5,                       # Rebel/Military involvement
  "cooperative" = 3                    # Domestic Private
)

# Function to map ownership type to score
get_ownership_score <- function(ownership) {
  score <- ownership_scores[ownership]
  if (is.null(score)) {
    return(NA)  # Handle any unexpected values
  } else {
    return(score)
  }
}

# Apply the function to create the ownership score column
grd <- grd %>%
  mutate(ownership_score = sapply(mineownership, get_ownership_score))

# Calculate the average ownership score for each PRIO grid cell and year
aggregate_scores <- grd %>%
  group_by(gid, year) %>%
  summarise(avg_ownership_score = mean(ownership_score, na.rm = TRUE), .groups = 'drop')

# Check the results
print(head(aggregate_scores))
```

