---
title: "Replicating Dube Vargas (baseline)"
author: "Christian"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Replicating Dube Vargas As Baseline
Note: Dube and Vargas look at the period 1988 to 2005. I have data from 1996 to 2014.
They use average number of crude oil barrels per day in 1988, the beginning of their sample period. Since I don't have resource data by day but instead by year, I could divide each by 365 to get average daily production? Since my resource dataset covers 1994 and 1995 and I could measure oil production based on those years, i.e. before the conflict period (or copy Dube & Vargas and use their 1988 measure?)

# Import the Colombia Conflict Data from xSub
I import the Colombia conflict data from xSub with the following specifications:
* multi-source 1km - 1day filter, undirected dyad, with unit of analysis Admin-2 (district, county) and temporal unit year.

```{r}
library(tidyverse)
library(haven)
library(here)

# Import the Colombia conflict data
conflict_colombia <- read_csv(here("Data", "xSub Conflict Data", "xSub_MELTT1km1dB_COL_adm2_year.csv"))

```

Let's check the dataset

# Getting the GRD Resources Data: Filter for Colombia

```{r}
# Load packages
library(tidyverse)
library(haven)
library(here)

# Import dta file
grd <- read_dta(here("Data", "GRD Dataset", "dfhsw_GRD_public_v1.dta"))
```

## Filter for country and data range
Now let's filter for Colombia and the year range matching the conflict dataset. We also drop some irrelevant columns. I include 1994 to form the baseline lootability estimate (before sample period starting 1995). 1994 is the earliest year in the dataset for Colombia.

```{r}
# Filter for Colombia
colombia_grd <- grd |> 
  filter(country == "colombia") |> 
  filter(year >= 1994 & year <= 2014) |> 
  select(-c("COWcode", "wb_ccode", "gwno", "log_comtrade_val", "log_wb_val", "log_usgs_val", "log_multicolour_val"))

colombia_grd |>
  head()
```

Let's check the summary statistics for the filtered dataset. First, let's list all the variables we have.

```{r}
# List all variables
colombia_grd |> 
  names()

```

First, what is the date coverage?

```{r}
# Date coverage
colombia_grd |> 
  select(year) |> 
  distinct() |> 
  pull(year)
```

So the data covers the years 1994 to 2014.

How many different resources are included?

```{r}
# Distinct types of resources, print them
colombia_grd |> 
  select(resource) |> 
  distinct() |> 
  pull(resource)

```
# Clean admin2 data

```{r}
# First display how many unique values there are of admin1
unique_admin2 <- colombia_grd |> 
  select(admin2) |> 
  distinct() |> 
  nrow()

print(paste("Unique admin2 values:", unique_admin2))

# Print the names of admin1 unique values
unique_admin2_names <- colombia_grd |> 
  select(admin2) |> 
  distinct() |> 
  pull()

print(unique_admin2_names)
```

