---
title: "Conflict Visualisations xSub"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import xSub Data

```{r}
library(xSub)
library(sf)
library(ggplot2)
library(dplyr)
library(xSub)
library(here)
library(haven)
library(stringr)

```

```{r}
# Load the PRIO grid cell shapefile
priogrid <- st_read(here("Data", "PRIO Shape Files", "priogrid_cell.shp"))

# Load the world administrative boundaries shapefile
world <- st_read(here("Data", "PRIO Shape Files", "world-administrative-boundaries.shp"))
```


```{r}
panel_data <- read_dta(here::here("..", "Stata Regressions", "code_tables_results", "Data", "matched_grd_xsub_withlags.dta"))

```

```{r}
# Subset PRIO-GRID shapefile to match the spatial extent of xSub data
priogrid <- priogrid %>%
  filter(gid %in% panel_data$gid)

```

```{r}
# Create world map showing conflict counts in 1994
# Extract xSub data for the year 2015
panel_data_1994 <- panel_data %>% 
  filter(year == 1994 & country == "colombia")
  

# Drop overlapping columns except for the column used for merging (gid)
panel_data_1994 <- panel_data_1994 %>% 
  select(gid, setdiff(names(panel_data_1994), names(priogrid)))

# Merge PRIO-GRID shapefile with xSub data
map_2015 <- merge(priogrid, panel_data_1994, by.x = "gid", by.y = "gid")

# Plot the data using ggplot2
ggplot(data = st_as_sf(map_2015)) +
  geom_sf(aes(fill = ACTION_ANY_xS)) +
  labs(title = "Conflict in PRIO-GRID cells (2015)", fill = "Actions") +
  theme_minimal()

```

```{r}
# Filter the panel data to include only Colombia
panel_data_col <- panel_data %>% filter(country == "colombia")

# Group by gid and calculate the average of ACTION_ANY_xS over all years
average_conflict <- panel_data_col %>%
  group_by(gid) %>%
  summarize(avg_ACTION_ANY_xS = mean(ACTION_ANY_xS, na.rm = TRUE))

# Check the aggregated data
print("Aggregated data (average ACTION_ANY_xS):")
print(head(average_conflict))

# Merge PRIO-GRID shapefile with the aggregated data
map_col_avg <- merge(priogrid, average_conflict, by = "gid")

# Check the merged data
print("Merged data with average conflict counts:")
print(summary(map_col_avg))

# Plot the data using ggplot2
ggplot(data = st_as_sf(map_col_avg)) +
  geom_sf(aes(fill = avg_ACTION_ANY_xS)) +
  labs(title = "Average Conflict Counts in PRIO-GRID cells - Colombia", fill = "Average Actions") +
  theme_minimal()
```

```{r}
# Filter the panel data to include only countries in the Americas
# Assuming there's a 'continent' column in the panel_data
panel_data_americas <- panel_data %>% filter(continent == "americas")

# Group by gid and calculate the average of ACTION_ANY_xS over all years
average_conflict_americas <- panel_data_americas %>%
  group_by(gid) %>%
  summarize(avg_ACTION_ANY_xS = mean(ACTION_ANY_xS, na.rm = TRUE))

# Check the aggregated data
print("Aggregated data (average ACTION_ANY_xS) for the Americas:")
print(head(average_conflict_americas))

# Merge PRIO-GRID shapefile with the aggregated data
map_americas_avg <- merge(priogrid, average_conflict_americas, by = "gid")

# Check the merged data
print("Merged data with average conflict counts for the Americas:")
print(summary(map_americas_avg))

# Create the plot using ggplot2
plot <- ggplot(data = st_as_sf(map_americas_avg)) +
  geom_sf(aes(fill = avg_ACTION_ANY_xS)) +
  scale_fill_viridis_c(option = "viridis") +
  labs(title = "Average Conflict Counts in PRIO-GRID cells - Americas", fill = "Average Actions") +
  theme_minimal()

# Save the plot with specified dimensions and resolution
ggsave("average_conflict_counts_americas.png", plot, width = 12, height = 8, dpi = 300)
```
```{r}
unique_countries_americas <- panel_data_americas %>% select(country) %>% distinct()

```

# Import pure xSub Data

```{r}
# Clear workspace and free up unused memory
rm(list = ls())

library(xSub)
library(sf)
library(ggplot2)
library(dplyr)
library(xSub)
library(here)
library(haven)
library(stringr)
library(readr)
library(viridis)

# Load 1k1d xSub data
xsub_data <- read_csv(here("Data", "xSub Conflict Data", "xsub_1k1d_prio_year_undirected.csv"))

xsub_data <- xsub_data |> 
  filter(YEAR >= 1994 & YEAR <= 2014) |> 
  rename(gid = PRIO_GID)
```

```{r}
# Load the PRIO grid cell shapefile
priogrid <- st_read(here("Data", "PRIO Shape Files", "priogrid_cell.shp"))

# Load the world administrative boundaries shapefile
world <- st_read(here("Data", "PRIO Shape Files", "world-administrative-boundaries.shp"))
```

```{r}
# Select specific columns from the world shapefile including geometry
world_selected <- world %>%
  select(iso3, color_code, continent, region, geometry)

# Merge the selected columns from the shapefile with the xsub_data using the ISO and iso3 columns
merged_data_xsub <- xsub_data %>%
  left_join(world_selected, by = c("ISO" = "iso3"), relationship = "many-to-many")

```


```{r}
# Define list of countries
iso3_countries <- c(
  "AFG", "AGO", "ALB", "ARE", "ARG", "AUS", "ARM", "AZE", "BDI", "BEL", "BEN", "BFA", "BGD", "BGR", "BHR", "BHS", "BIH", "BLR", "BLZ", "BOL", "BRA", "BRB", "BTN", "BWA", "CAF", "CAN", "CHE", "CHN", "CIV", "CMR", "COD", "COG", "COL", "COM", "CRI", "CUB", 
  "CYP", "DEU", "DJI", "DOM", "DZA", "ECU", "EGY", "ERI", "ESP", "EST", "ETH", "FIN", "FRA", "GAB", "GBR", "GEO", "GHA", "GIB", "GIN", "GMB", "GNB", "GNQ", "GRC", "GTM", "GUY", "HND", "HRV", "HTI", "IDN", "IND", "IRL", "IRN", "IRQ",
  "ISR", "ITA", "JAM", "JOR", "JPN", "KAZ", "KEN", "KGZ", "KHM", "KWT", "LAO", "LBN", "LBR", "LBY", "LKA", "LSO", "LTU", "LVA", "MAR", "MDA", "MDG", "MDV", "MEX", "MKD", "MLI", "MLT", "MMR", "MNE",
  "MOZ", "MRT", "MUS", "MWI", "MYS", "NAM", "NER", "NGA", "NIC", "NLD", "NOR", "NPL", "NZL", "OMN", "PAK", "PAN", "PER", "PHL", "PNG", "PRI", "PRT", "PRY", "PSE", "QAT", "ROU", "RUS", "RWA", "SAU", "SDN", "SEN", "SLB", "SLE", "SLV", "SOM",
  "SRB", "SSD", "SVK", "SWE", "SWZ", "SYR", "TCD", "TGO", "THA", "TJK", "TKM", "TLS", "TTO", "TUN", "TUR", "TZA", "UGA", "UKR", "USA", "UZB", "VEN", "VNM", "YEM", "ZAF", "ZMB", "ZWE"
)

# Check which ISO values are not in the xSub data
missing_iso_values <- setdiff(iso3_countries, xsub_data$ISO)

# Print the missing ISO values
print("Missing ISO values:")
print(missing_iso_values)
```
```{r}
# Check if I can get these extra xSub data separately
# Loop through each ISO3 country
SWE_data <- get_xSub_multi(sources_type = "multiple",
                                 data_type = "panel",
                                 country = "SWE",
                                 space_unit = "priogrid",
                                 time_unit = "year",
                                 geo_window = "1 km",
                                 time_window = "1 day",
                                 dyad_type = "undirected",
                                 merge_files = TRUE,
                                 out_dir =  "D:/Dropbox/CHRISTIAN/MPhil/Thesis/R Code/Data",
                                 write_file = FALSE,  # Set to FALSE to manage data within R
                                 write_format = "csv",
                                 verbose = TRUE)
```
When downloading Argentina, I only get observations from 1992 using 1k1d undirected.
When downloading Australia, I only get observations from 2018 using 1k1d undirected.
When downloading Cyprus, I only get observations from 2018 and 2019 using 1k1kd undirected
When downloading Estonia, I only get observations from 1987-1992 using 1k1d undirected.
When downloading Georgia, I get observations from 1987-2019 using 1k1d undirected.THIS SHOULD HAVE BEEN INCLUDED! BAD CODING FROM XSUB!!!
When downloading Ghana, this is not possible
When downloading Gibraltar I get one observation from 1988 using 1k1d undirected
When downloading Guinea I get observations from 1993-2019 using 1k1d undirected - bad coding from xSub again.
When downloading Maldives, this is not possible
When downloading Montenegro, I only get observations from 2018-2019
When downloading New Zealand, I only get observations from 1974
When downloading Puerto Rico, this is not possible.
When downloading Sweden, this is not possible.

In conclusion, Guinea and Georgia should have been included in the xSub data. The rest of the countries are not possible to include in the xSub data.



CONTINUE WITH THE REST OF THE COUNTRIES


```{r}
# Select specific columns from the world shapefile including geometry
world_selected <- world %>%
  select(iso3, color_code, continent, region, geometry)

# Merge the selected columns from the shapefile with the GRD data using the wb_ccode and iso3 columns
merged_data <- xsub_data %>%
  left_join(world_selected, by = c("ISO" = "iso3"))
```

```{r}
# Create a named vector for ISO to continent mapping
iso_to_continent <- c(
  "AFG" = "Asia", "AGO" = "Africa", "ALB" = "Europe", "ARE" = "Asia", "AUS" = "Oceania",
  "ARM" = "Asia", "AZE" = "Asia", "BDI" = "Africa", "BEL" = "Europe", "BEN" = "Africa",
  "BFA" = "Africa", "BGD" = "Asia", "BGR" = "Europe", "BHR" = "Asia", "BHS" = "Americas",
  "BIH" = "Europe", "BLR" = "Europe", "BLZ" = "Americas", "BOL" = "Americas", "BRA" = "Americas",
  "BRB" = "Americas", "BTN" = "Asia", "BWA" = "Africa", "CAF" = "Africa", "CAN" = "Americas",
  "CHE" = "Europe", "CHN" = "Asia", "CIV" = "Africa", "CMR" = "Africa", "COD" = "Africa",
  "COG" = "Africa", "COL" = "Americas", "COM" = "Africa", "CRI" = "Americas", "CUB" = "Americas",
  "CYP" = "Europe", "DEU" = "Europe", "DJI" = "Africa", "DOM" = "Americas", "DZA" = "Africa",
  "ECU" = "Americas", "EGY" = "Africa", "ERI" = "Africa", "ESP" = "Europe", "EST" = "Europe",
  "ETH" = "Africa", "FIN" = "Europe", "FRA" = "Europe", "GAB" = "Africa", "GBR" = "Europe",
  "GIB" = "Europe", "GMB" = "Africa", "GNB" = "Africa", "GNQ" = "Africa", "GRC" = "Europe",
  "GTM" = "Americas", "GUY" = "Americas", "HND" = "Americas", "HRV" = "Europe", "HTI" = "Americas",
  "IDN" = "Asia", "IND" = "Asia", "IRL" = "Europe", "IRN" = "Asia", "IRQ" = "Asia",
  "ISR" = "Asia", "ITA" = "Europe", "JAM" = "Americas", "JOR" = "Asia", "JPN" = "Asia",
  "KAZ" = "Asia", "KEN" = "Africa", "KGZ" = "Asia", "KHM" = "Asia", "KWT" = "Asia",
  "LAO" = "Asia", "LBN" = "Asia", "LBR" = "Africa", "LBY" = "Africa", "LKA" = "Asia",
  "LSO" = "Africa", "LTU" = "Europe", "LVA" = "Europe", "MAR" = "Africa", "MDA" = "Europe",
  "MDG" = "Africa", "MEX" = "Americas", "MKD" = "Europe", "MLI" = "Africa", "MLT" = "Europe",
  "MMR" = "Asia", "MNE" = "Europe", "MOZ" = "Africa", "MRT" = "Africa", "MUS" = "Africa",
  "MWI" = "Africa", "MYS" = "Asia", "NAM" = "Africa", "NER" = "Africa", "NGA" = "Africa",
  "NIC" = "Americas", "NLD" = "Europe", "NOR" = "Europe", "NPL" = "Asia", "NZL" = "Oceania",
  "OMN" = "Asia", "PAK" = "Asia", "PAN" = "Americas", "PER" = "Americas", "PHL" = "Asia",
  "PNG" = "Oceania", "PRT" = "Europe", "PRY" = "Americas", "PSE" = "Asia", "QAT" = "Asia",
  "ROU" = "Europe", "RUS" = "Europe", "RWA" = "Africa", "SAU" = "Asia", "SDN" = "Africa",
  "SEN" = "Africa", "SLB" = "Oceania", "SLE" = "Africa", "SLV" = "Americas", "SOM" = "Africa",
  "SRB" = "Europe", "SSD" = "Africa", "SVK" = "Europe", "SWE" = "Europe", "SWZ" = "Africa",
  "SYR" = "Asia", "TCD" = "Africa", "TGO" = "Africa", "THA" = "Asia", "TJK" = "Asia",
  "TKM" = "Asia", "TLS" = "Asia", "TTO" = "Americas", "TUN" = "Africa", "TUR" = "Asia",
  "TZA" = "Africa", "UGA" = "Africa", "UKR" = "Europe", "USA" = "Americas", "UZB" = "Asia",
  "VEN" = "Americas", "VNM" = "Asia", "YEM" = "Asia", "ZAF" = "Africa", "ZMB" = "Africa",
  "ZWE" = "Africa", "NA" = "none"  # Assign "none" for NA entries
)

# Create the continent_xS variable using the named vector
xsub_data <- xsub_data %>%
  mutate(continent_xS = recode(ISO, !!!iso_to_continent)) |> 
  select(SOURCE, continent_xS, everything())

# Print the first few rows to verify
print("Panel data with continent_xS variable:")
print(head(xsub_data))

```



## Europe

```{r}
panel_data_europe <- xsub_data |> 
  filter(continent_xS == "Europe")
```


```{r}
# Calculate the average ACTION_ANY values across years for each grid cell (gid)
average_conflict_europe <- panel_data_europe %>%
  group_by(gid) %>%
  summarize(avg_ACTION_ANY = mean(ACTION_ANY, na.rm = TRUE))

# Merge PRIO-GRID shapefile with the aggregated data
map_europe_avg <- merge(priogrid, average_conflict_europe, by = "gid")

```


```{r}
# Display all distinct years
distinct_years <- xsub_data %>% select(YEAR) %>% distinct()
```


```{r}
# Calculate the average ACTION_ANY values across years for each grid cell (gid)
# Add a smaller constant (0.1) to avoid taking the log of zero
average_conflict_europe <- panel_data_europe %>%
  group_by(gid) %>%
  summarize(avg_ACTION_ANY = mean(ACTION_ANY, na.rm = TRUE) + 0.1)

# Merge PRIO-GRID shapefile with the aggregated data
map_europe_avg <- merge(priogrid, average_conflict_europe, by = "gid")

# Create the plot using ggplot2 with a log transformation and simplified breaks
plot <- ggplot(data = st_as_sf(map_europe_avg)) +
  geom_sf(aes(fill = avg_ACTION_ANY), color = NA) +
  scale_fill_viridis_c(
    option = "plasma",
    trans = "log10",
    breaks = c(0.1, 1, 10, 100, 1000),
    labels = c("0", "1", "10", "100", "1000")
  ) +
  labs(title = "Average Conflict In Cells (1994-2014): Europe", fill = "Average Action (ANY) Count (Log Scale)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 30),
    legend.position = "right",
    legend.text = element_text(size = 22),
    legend.title = element_text(size = 24),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  ) +
  coord_sf(crs = st_crs(3035))  # European projection

# Save the plot as PDF with specified dimensions and resolution
ggsave("average_conflict_counts_europe_log_scale.pdf", plot, width = 20, height = 12, units = "in")

# Print the plot
print(plot)
```

## Africa

```{r}
# Filter the panel data to include only Africa
panel_data_africa <- xsub_data %>%
  filter(continent_xS == "Africa")

# Calculate the average ACTION_ANY values across years for each grid cell (gid)
# Add a smaller constant (0.1) to avoid taking the log of zero
average_conflict_africa <- panel_data_africa %>%
  group_by(gid) %>%
  summarize(avg_ACTION_ANY = mean(ACTION_ANY, na.rm = TRUE) + 0.1)

# Merge PRIO-GRID shapefile with the aggregated data
map_africa_avg <- merge(priogrid, average_conflict_africa, by = "gid")

# Create the plot using ggplot2 with a log transformation and simplified breaks
plot_africa <- ggplot(data = st_as_sf(map_africa_avg)) +
  geom_sf(aes(fill = avg_ACTION_ANY), color = NA) +
  scale_fill_viridis_c(
    option = "plasma",
    trans = "log10",
    breaks = c(0.1, 1, 10, 100, 1000),
    labels = c("0", "1", "10", "100", "1000")
  ) +
  labs(title = "Average Conflict In Cells (1994-2014): Africa", fill = "Average Action (ANY) Count (Log Scale)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 30),
    plot.title.position = "plot",
    legend.position = "right",
    legend.text = element_text(size = 22),
    legend.title = element_text(size = 24),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  ) +
  coord_sf(crs = st_crs(3035))  # Suitable projection

# Save the plot as a larger PDF with specified dimensions and resolution
ggsave("average_conflict_counts_africa_log_scale.pdf", plot_africa, width = 20, height = 12, units = "in")

# Print the plot
print(plot_africa)
```

## Asia

```{r}
# Filter the panel data to include only Africa
panel_data_asia <- xsub_data %>%
  filter(continent_xS == "Asia")

# Calculate the average ACTION_ANY values across years for each grid cell (gid)
# Add a smaller constant (0.1) to avoid taking the log of zero
average_conflict_asia <- panel_data_asia %>%
  group_by(gid) %>%
  summarize(avg_ACTION_ANY = mean(ACTION_ANY, na.rm = TRUE) + 0.1)

# Merge PRIO-GRID shapefile with the aggregated data
map_asia_avg <- merge(priogrid, average_conflict_asia, by = "gid")

# Create the plot using ggplot2 with a log transformation and simplified breaks
plot_asia <- ggplot(data = st_as_sf(map_asia_avg)) +
  geom_sf(aes(fill = avg_ACTION_ANY), color = NA) +
  scale_fill_viridis_c(
    option = "plasma",
    trans = "log10",
    breaks = c(0.1, 1, 10, 100, 1000),
    labels = c("0", "1", "10", "100", "1000")
  ) +
  labs(title = "Average Conflict In Cells (1994-2014): Asia", fill = "Average Action (ANY) Count (Log Scale)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 30),
    plot.title.position = "plot",
    legend.position = "right",
    legend.text = element_text(size = 22),
    legend.title = element_text(size = 24),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  ) +
  coord_sf(crs = "+proj=wintri")  # Suitable projection

# Save the plot as a larger PDF with specified dimensions and resolution
ggsave("average_conflict_counts_asia_log_scale.pdf", plot_asia, width = 20, height = 12, units = "in")

# Print the plot
print(plot_asia)
```

## Americas

```{r}
# Filter the panel data to include only Americas
panel_data_americas <- xsub_data %>%
  filter(continent_xS == "Americas")

# Calculate the average ACTION_ANY values across years for each grid cell (gid)
# Add a smaller constant (0.1) to avoid taking the log of zero
average_conflict_americas <- panel_data_americas %>%
  group_by(gid) %>%
  summarize(avg_ACTION_ANY = mean(ACTION_ANY, na.rm = TRUE) + 0.1)

# Merge PRIO-GRID shapefile with the aggregated data
map_americas_avg <- merge(priogrid, average_conflict_americas, by = "gid")

# Create the plot using ggplot2 with a log transformation and simplified breaks
plot_americas <- ggplot(data = map_americas_avg) +
  geom_sf(aes(fill = avg_ACTION_ANY), color = NA) +
  scale_fill_viridis_c(
    option = "plasma",
    trans = "log10",
    breaks = c(0.1, 1, 10, 100, 1000),
    labels = c("0", "1", "10", "100", "1000")
  ) +
  labs(title = "Average Conflict In Cells (1994-2014): Americas", fill = "Average Action (ANY) Count (Log Scale)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 30),
    plot.title.position = "plot",
    legend.position = "right",
    legend.text = element_text(size = 22),
    legend.title = element_text(size = 24),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  ) +
  coord_sf(crs = st_crs(5070))  # Albers Equal-Area Conic projection using EPSG code

# Save the plot as a larger PDF with specified dimensions and resolution
ggsave("average_conflict_counts_americas_log_scale.pdf", plot_americas, width = 20, height = 12, units = "in")

# Print the plot
print(plot_americas)
```

## Oceania

```{r}
# Filter the panel data to include only Oceania
panel_data_oceania <- xsub_data %>%
  filter(continent_xS == "Oceania")

# Calculate the average ACTION_ANY values across years for each grid cell (gid)
# Add a smaller constant (0.1) to avoid taking the log of zero
average_conflict_oceania <- panel_data_oceania %>%
  group_by(gid) %>%
  summarize(avg_ACTION_ANY = mean(ACTION_ANY, na.rm = TRUE) + 0.1)

# Merge PRIO-GRID shapefile with the aggregated data
map_oceania_avg <- merge(priogrid, average_conflict_oceania, by = "gid")

# Apply the CRS to the merged data using the Pacific-centered Lambert Azimuthal Equal-Area projection
map_oceania_avg <- st_transform(st_as_sf(map_oceania_avg), crs = "+proj=laea +lon_0=160 +lat_0=-10 +datum=WGS84 +units=m +no_defs")

# Create the plot using ggplot2 with a log transformation and simplified breaks
plot_oceania <- ggplot(data = map_oceania_avg) +
  geom_sf(aes(fill = avg_ACTION_ANY), color = NA) +
  scale_fill_viridis_c(
    option = "plasma",
    trans = "log10",
    breaks = c(0.1, 1, 10, 100, 1000),
    labels = c("0", "1", "10", "100", "1000")
  ) +
  labs(title = "Average Conflict In Cells (1994-2014): Oceania", fill = "Average Action (ANY) Count (Log Scale)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 30),
    plot.title.position = "plot",
    legend.position = "right",
    legend.text = element_text(size = 22),
    legend.title = element_text(size = 24),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  ) +
  coord_sf(crs = st_crs("+proj=laea +lon_0=160 +lat_0=-10 +datum=WGS84 +units=m +no_defs"))  # Pacific-centered Lambert Azimuthal Equal-Area projection

# Save the plot as a larger PDF with specified dimensions and resolution
ggsave("average_conflict_counts_oceania_log_scale.pdf", plot_oceania, width = 20, height = 12, units = "in")

# Print the plot
print(plot_oceania)

```

## World Map

```{r}
# Calculate the average ACTION_ANY values across years for each grid cell (gid)
# Add a smaller constant (0.1) to avoid taking the log of zero
average_conflict_world <- xsub_data %>%
  group_by(gid) %>%
  summarize(avg_ACTION_ANY = mean(ACTION_ANY, na.rm = TRUE) + 0.1)

# Merge PRIO-GRID shapefile with the aggregated data
map_world_avg <- merge(priogrid, average_conflict_world, by = "gid")

# Apply the CRS to the merged data using the Robinson projection (ESRI:54030)
map_world_avg <- st_transform(st_as_sf(map_world_avg), crs = st_crs("ESRI:54030"))

# Create the plot using ggplot2 with a log transformation and simplified breaks
plot_world <- ggplot(data = map_world_avg) +
  geom_sf(aes(fill = avg_ACTION_ANY), color = NA) +
  scale_fill_viridis_c(
    option = "plasma",
    trans = "log10",
    breaks = c(0.1, 1, 10, 100, 1000),
    labels = c("0", "1", "10", "100", "1000")
  ) +
  labs(title = "Average Conflict In Cells (1994-2014): World", fill = "Average Action (ANY) Count (Log Scale)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 30),
    plot.title.position = "plot",
    legend.position = "right",
    legend.text = element_text(size = 22),
    legend.title = element_text(size = 24),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  ) +
  coord_sf(crs = st_crs("ESRI:54030"))  # Robinson projection using ESRI code

# Save the plot as a larger PDF with specified dimensions and resolution
ggsave("average_conflict_counts_world_log_scale.pdf", plot_world, width = 20, height = 12, units = "in")

# Print the plot
print(plot_world)

```

# Create Summary Table For xSub Data

```{r}
library(dplyr)
library(haven)
library(here)
library(knitr)
library(kableExtra)
library(stringr)

summary_table_xsub <- xsub_data %>%
  select(continent_xS, NAME_0, gid, YEAR) %>%
  group_by(continent_xS, NAME_0) %>%
  summarize(
    `Distinct Prio-Grid Cells` = n_distinct(gid),
    `Years Covered` = paste(min(YEAR), "-", max(YEAR)),
    .groups = 'drop'
  )
# Calculate totals
total_countries <- n_distinct(summary_table_xsub$NAME_0)
total_priogrid_cells <- sum(summary_table_xsub$`Distinct Prio-Grid Cells`)

# Add the total row
summary_table_xsub <- summary_table_xsub %>%
  bind_rows(tibble(
    continent_xS = "Total",
    NAME_0 = "Total",
    `Distinct Prio-Grid Cells` = total_priogrid_cells,
    `Years Covered` = ""
  ))

# Convert the summary table to LaTeX format
latex_table_xsub <- kable(summary_table_xsub, format = "latex", booktabs = TRUE, longtable = TRUE, 
                          caption = "Summary of Distinct PRIO-GRID Cells and Years by Country and Continent",
                          col.names = c("Continent", "Country", "Distinct Prio-Grid Cells", "Years Covered")) %>%
  kable_styling(latex_options = c("striped", "hold_position", "repeat_header")) %>%
  pack_rows(index = table(summary_table_xsub$continent_xS)) %>%
  kable_styling(font_size = 8)  # Reduce the font size

# Define the file path using the here package
tex_file_xsub <- here::here("RTables", "summary_table_continent_country_priogrid_years_xsub.tex")

# Write the LaTeX table to a file
writeLines(latex_table_xsub, tex_file_xsub)

```
# Use World Shape File

```{r}
library(dplyr)
library(haven)
library(here)
library(knitr)
library(kableExtra)
library(stringr)

# Create a summary table with distinct PRIO-GRID cells and year range for each country, grouped by continent
summary_table_xsub <- merged_data_xsub %>%
  select(continent, NAME_0, gid, YEAR) %>%
  mutate(NAME_0 = str_to_title(NAME_0)) %>%
  group_by(continent, NAME_0) %>%
  summarize(
    `Distinct Prio-Grid Cells` = n_distinct(gid),
    `Years Covered` = paste(min(YEAR), "-", max(YEAR)),
    .groups = 'drop'
  )

# Convert the summary table to LaTeX format
latex_table_xsub <- kable(summary_table_xsub, format = "latex", booktabs = TRUE, longtable = TRUE, 
                          caption = "Summary of Distinct PRIO-GRID Cells and Years by Country and Continent",
                          col.names = c("Continent", "Country", "Distinct Prio-Grid Cells", "Years Covered")) %>%
  kable_styling(latex_options = c("striped", "hold_position", "repeat_header")) %>%
  pack_rows(index = table(summary_table_xsub$continent)) %>%
  kable_styling(font_size = 8)  # Reduce the font size

# Define the file path using the here package
tex_file_xsub <- here::here("RTables", "summary_table_continent_country_priogrid_years_xsub.tex")

# Write the LaTeX table to a file
writeLines(latex_table_xsub, tex_file_xsub)

```

```{r}
# Show rows that have continent Antarctica
merged_data_xsub %>%
  filter(continent == "Antarctica")
```

